---
title: "BuoyDataScraping"
author: "SM"
date: "11/14/2015"
output: html_document
---

http://www.ndbc.noaa.gov/historical_data.shtml


I pulled data from url.  It was already a data frame and looked nice, so just added the variable names to the top.  The class of some of the columns were integer, so I just changed all columns to be of class numeric.

The data was already pretty tidy.


Note: 
-we probably want to make year month day into ymd.  Or maybe make year month day hour min all one variable?  
-Would we need to use lubridate?  
-we would also want to pull out the variable we want   
-we would want to set up a naming convention for all the buoy datasets  


```{r}


require(lubridate)
require(dplyr)
require(tidyr)

```



```{r}
#make a loop to pull in all buoy data for a set of buoys
#41012h2004
buoyNumber <- c( 41012,42007)
buoyYear <- c(2004,2009)
setOfBuoys <- vector("list",length(buoyNumber))

dataName <- seq(1,length(buoyNumber), by=1)

for(j in 1:length(buoyNumber)){
  site <- paste("http://www.ndbc.noaa.gov/view_text_file.php?filename=", buoyNumber[j], "h", buoyYear[j], ".txt.gz&dir=data/historical/stdmet/", sep ="")

  # call in data with try command in while loop
  i <- 1
  while (i < 2){
    data <- try(read.table(site,sep="", fill=TRUE, header=TRUE, comment.char=""))
    if (class(data) == "try-error") {
      next
    } else {
      i <- i + 1
    }
  }

#if buoy data has comment sign (#) in variable names
if(length(grep("X.",names(data)))>0){
  #ie names variable X.YY to YY
  names(data)[grep("X.",names(data))] <-
    substr(names(data)[grep("X.",names(data))],                                                start=3,nchar(names(data)[grep("X.",names(data))]))
  #removes row of units
  data <- data[-1,]
  
  #change YY to YYYY, WDIR to WD, PRES to BAR
  names(data)[grep("YY", names(data))] <- "YYYY"
  names(data)[grep("WDIR", names(data))] <- "WD"
  names(data)[grep("PRES", names(data))] <- "BAR"
  
  #remove mm (min col)
  data <- data[,-grep("mm", names(data))]
}
  
data <- as.data.frame(apply(data[,1:(length(data))], 2, as.numeric))
  
#make new variable ymd_hms w lubridate (set all min sec to 00)
data<-data%>%mutate(ymdday=ymd_hms(paste(YYYY,"-", MM,"-",DD, "-", hh, "-", 00, "-", 00)))
  
#make all columns numeric and put in vector
setOfBuoys[[j]]<- as.data.frame(data)

}

```

The code above uses a for loop to pull in dataset for a set of buoys.  It stores the dataset for each buoy in a list called setOf Buoys.  In order to run this for several buoys we will need to add another for loop and pick a naming convention for setOfBuoys.


NEED TO TEST CODE ABOVE
PROBLEM THAT NOT ALL BUOY DATASETS HAVE SAME NUMBER OF VARIABLES