---
title: "idea for data scraping for buoys"
author: "SM"
date: "11/6/2015"
output: html_document
---

Note: Starting a new file for actual buoy data scraping called BuoyDataScraping

http://www.ndbc.noaa.gov/historical_data.shtml

```{r}
#example data: 2009 buoy 32487
#the web url
site <- "http://www.ndbc.noaa.gov/view_text_file.php?filename=32487h2009.txt.gz&dir=data/historical/stdmet/"

# call in data with try command in while loop
i <- 1
while (i < 2){
  theData <- try(read.table(site,sep="",comment.char=""))
  if (class(theData) == "try-error") {
    next
  } else {
    i <- i + 1
  }
}

#look at the data
head(theData, 20)

#data is already a data frame
class(theData)

#assign to new data frame to do cleaning on
cleanData<- theData

#add names to variables
names(cleanData)<-c("year", "month", "day", "hour", "min", "WDIR", "WSPD", "GST", "WVHT", "DPD", "APD", "MWD", "PRES", "ATMP", "WTMP", "DEWP", "VIS", "TIDE")

#FOR .TXT THAT HAVE 2 ROWS OF COMMENTED OUT VARIABLE NAMES
#ADD A CHECK TO SEE IF FIRST COL OF VARIABLE NAMES MATCHES NAMES ASSIGNED
#WILL ALSO NEED TO REMOVE FIRST 2 COLS BEFORE MAKING NUMERIC

#make all columns numeric
cleanData <- as.data.frame(apply(cleanData[,1:length(cleanData)], 2, as.numeric))

#check out clean data
head(cleanData)
```

I pulled data from url.  It was already a data frame and looked nice, so just added the variable names to the top.  The class of some of the columns were integer, so I just changed all columns to be of class numeric.

The data was already pretty tidy.


Note: 
-we probably want to make year month day into ymd.  Or maybe make year month day hour min all one variable?  
-Would we need to use lubridate?  
-we would also want to pull out the variable we want   
-we would want to set up a naming convention for all the buoy datasets  

```{r}
#make a loop to pull in all buoy data for a set of buoys

buoyNumber <- c(32487,41024)
buoyYear <- c(2009,2014)
setOfBuoys <- vector("list",length(buoyNumber))

dataName <- seq(1,length(buoyNumber), by=1)

for(j in 1:length(buoyNumber)){
  site <- paste("http://www.ndbc.noaa.gov/view_text_file.php?filename=", buoyNumber[j], "h", buoyYear[j], ".txt.gz&dir=data/historical/stdmet/", sep ="")

  # call in data with try command in while loop
  i <- 1
  while (i < 2){
    data <- try(read.table(site,sep="",comment.char=""))
    if (class(data) == "try-error") {
      next
    } else {
      i <- i + 1
    }
  }



  #add names to variables
  names(data)<-c("year", "month", "day", "hour", "min", "WDIR", "WSPD", "GST","WVHT","DPD", "APD", "MWD", "PRES", "ATMP", "WTMP", "DEWP", "VIS", "TIDE")
  
#FOR .TXT THAT HAVE 2 ROWS OF COMMENTED OUT VARIABLE NAMES
#ADD A CHECK TO SEE IF FIRST COL OF VARIABLE NAMES MATCHES NAMES ASSIGNED
#WILL ALSO NEED TO REMOVE FIRST 2 COLS BEFORE MAKING NUMERIC
  

#make all columns numeric and put in vector

setOfBuoys[[j]]<- as.data.frame(apply(data[,1:length(data)], 2, as.numeric))

}

```

The code above uses a for loop to pull in dataset for a set of buoys.  It stores the dataset for each buoy in a list called setOf Buoys.  In order to run this for several buoys we will need to add another for loop and pick a naming convention for setOfBuoys.


NEED TO TEST CODE ABOVE
PROBLEM THAT NOT ALL BUOY DATASETS HAVE SAME NUMBER OF VARIABLES