---
title: "Storms and Buoys Data"
author: "SM"
date: "11/14/2015"
output: html_document
---

Important Dataset:

finalStorms: has all of storm data 2000-2015, filtered out 1999 storms and made ymd_hms variable called time

buoyLocations: has all of the buoy numbers, locations (long,lat), and TRUE/FALSE for years 2000 to 2015 indicating whether it has historical data from that year

weightedMeans: has the weighed means of each buoy variables and the row number of the storm data that these weighted means refer to



```{r, message=FALSE}

require(dplyr)
require(tidyr)
require(geosphere)
require(lubridate)

```


Bring in all the storm data (2000 to 2015) and put in one dataset called finalStorms.
```{r}
#data from http://www.ncdc.noaa.gov/ibtracs/index.php?name=ibtracs-data
#ftp://eclipse.ncdc.noaa.gov/pub/ibtracs/v03r07/all/csv/year/


stormYearsFile <- c("2000.csv","2001.csv","2002.csv","2003.csv","2004.csv","2005.csv","2006.csv","2007.csv","2008.csv","2009.csv","2010.csv","2011.csv","2012.csv","2013.csv","2014.csv", "2015.csv")

allStorms <- vector("list",length(stormYearsFile)) 

#pull in csv files and assign to a list of datasets
for(i in 1:length(stormYearsFile)){
  
  allStorms[[i]] <-read.csv(paste("~/Comp-Stats-Project/storm",stormYearsFile[i], sep=""))
}

#only keep the first 19 columns
for(i in 1:length(allStorms)){
  allStorms[[i]] <- allStorms[[i]][,1:19]
}

#remove -999 from latitude and longitude
cleanStorms <- vector("list",length(stormYearsFile))

for(i in 1:length(allStorms)){
  cleanStorms[[i]] <- allStorms[[i]]%>% 
    filter(Latitude!=-999) %>% 
    filter(Longitude!=-999)
}

#check all storms datasets have same order of variables
checkingStorms <- c()
for(i in 2:length(cleanStorms)){
  checkingStorms[i-1] <- sum(names(cleanStorms[[1]]) == 
                             names(cleanStorms[[i]]))
}


#combine all storm datasets into 1 big dataset
finalStorms <- cleanStorms[[1]]
for(i in 2: length(cleanStorms)){
  finalStorms <- rbind(finalStorms, cleanStorms[[i]])
}

#add new column of time in ymd_hms type setupt using parse date time fuction
finalStorms <- finalStorms %>% mutate(time = parse_date_time(ISO_time, "%m%d%Y %H%M")) %>%
  filter(year(time) != 1999) # dont want storms in 2000 season that start in 1999

dput(finalStorms, "finalStorms.txt")

stormsNorthAtl <- finalStorms %>% 
  select(Basin, Serial_Num, Latitude, Longitude, time) %>%
  mutate(BasinChar = as.character(Basin))%>% filter(BasinChar == " NA") %>%
  mutate(Serial_NumChar = as.character(Serial_Num)) %>% 
  filter(Serial_NumChar %in% names(which(table(Serial_NumChar) > 1)))

dput(stormsNorthAtl, "stormsNorthAtl.txt")

```


This section below finds all the buoys for a row of storms data and calculates the weighted averages for each variable based on the distance of the buoy from the storm location.  We loop over the entire storm dataset going row by row.  The result is the dataset weightedMeans that has the average value of buoy variables calculated using weighted means and the rowNum in the storm data (finalStorms) it refers to.

```{r}

#table of buoy num, lat, long, and TRUE/FALSE if have data in years 2000 to 2015
buoyLocations <- dget("buoyLocsAndYears.txt") 

weightedMeans <- c()

#loops over every row in stormsNorthAtl dataset
#nrow(stormsNorthAtl)
for(k in 1:nrow(stormsNorthAtl)){
  isMatchingBuoysNew = FALSE
  
  firstRow <- stormsNorthAtl[k,]
  
  #column num that corresponds to variable yearX in buoyLocations file
  #yearX is the year of the storm for this row
  yearColNum <- which(names(buoyLocations)==paste("year", year(firstRow$time), sep=""))
  
  #only takes buoys w data from yearX
  buoyLocsThatYear <- buoyLocations %>% filter(buoyLocations[,yearColNum] == TRUE)
  
  stormCoords <- c(firstRow$Longitude, firstRow$Latitude)
  
    #function to compute dist from storm coord to buoy (buoy takes long/lat of buoy)
  computeDistanceTo <- function(buoy){
    return(distGeo(stormCoords,buoy))
  }
  
  #change 2e7 later
  #finds buoys within 2e7 distance to storm and get dataset of those buoys and their dists
  buoyDists <- buoyLocsThatYear %>% mutate(distToStorm = 
             apply(cbind(Longitude, Latitude), 1, computeDistanceTo)) %>%
    filter(distToStorm <= 500000) %>% select(BuoyNumber, distToStorm)
  

  buoyNumber <- buoyDists$BuoyNumber
  buoyYear <- year(firstRow$time)
  
  if(length(buoyNumber)!= 0){
    
    #dataset of buoys corresponding to storm row and their variables
    matchingBuoys <- as.data.frame(matrix(nrow = 1,ncol =19))
    names(matchingBuoys) = c("YYYY", "MM", "DD", "hh", "WD", "WSPD", "GST", "WVHT", "DPD", "APD", "MWD", "BAR", "ATMP" ,"WTMP","DEWP" ,"VIS", "TIDE", "ymdday", "BuoyNumber")
    
    #loops over buoys to pull in buoys and make dataset of buoys and their variables
    for(j in 1:length(buoyNumber)){
      site <- paste("http://www.ndbc.noaa.gov/view_text_file.php?filename=", buoyNumber[j], "h", buoyYear, ".txt.gz&dir=data/historical/stdmet/", sep ="")
    
      # call in buoy data with try command in while loop
      i <- 1
      while (i < 2){
        data <- try(read.table(site,sep="", fill=TRUE, header=TRUE, comment.char=""))
        if (class(data) == "try-error") {
          next
        } else {
          i <- i + 1
        }
      }
    
    #clean up if buoy data .txt has comment sign (#) -- 2007 and later
    if(length(grep("X.",names(data)))>0){
      #ie names variable X.YY to YY
      names(data)[grep("X.",names(data))] <-
        substr(names(data)[grep("X.",names(data))],                                                start=3,nchar(names(data)[grep("X.",names(data))]))
      #removes row of units
      data <- data[-1,]
      
      #change YY to YYYY, WDIR to WD, PRES to BAR
      names(data)[grep("YY", names(data))] <- "YYYY"
      names(data)[grep("WDIR", names(data))] <- "WD"
      names(data)[grep("PRES", names(data))] <- "BAR"
      
      #remove mm (min col)
      data <- data[,-grep("mm", names(data))]
    }
      
    #make everything numeric
    data <- as.data.frame(apply(data[,1:(length(data))], 2, as.numeric))
      
    #make new variable ymd_hms w lubridate (set all min 00 and sec to 00)
    nextRow<-data%>%mutate(ymdday=ymd_hms(paste(YYYY,"-", MM,"-",DD, "-", hh, "-", 00, "-", 00)))%>%
      filter(ymdday == firstRow$time) #only want data at time matching storm
    
    #if buoy data at matching time is not empty, add on buoyNumber 
    if(nrow(nextRow) > 0){
      #this storm row actually has data we want to use
      isMatchingBuoysNew = TRUE
      nextRow <- as.data.frame(c(nextRow, buoyNumber[j]))
      names(nextRow)[ncol(nextRow)] <- "BuoyNumber"
    
    
    
    #assign first row to empty matchingBuoys, then rbind on other nextRows sequentially
    if(j==1){
      matchingBuoys <- nextRow
    } else {
      matchingBuoys <- rbind(matchingBuoys, nextRow)
    }
      
    }
   
    }
    
    
    if(isMatchingBuoysNew){
    #join buoy data for that row with distances
    matchingBuoys <- matchingBuoys %>% 
      mutate(BuoyNumber = as.character(BuoyNumber))%>%
      left_join(buoyDists, by = "BuoyNumber")
    
    #converting 999 and 99 to NA
    matchingBuoys[,c(6:10, 16:17)][matchingBuoys[,c(6:10, 16:17)] == 99] <- NA
    
    matchingBuoys[,c(5, 11, 13:15)][matchingBuoys[,c(5, 11, 13:15)] == 999] <- NA
    
    
    # average repeated rows, with na.rm = T
    matchingBuoys <- matchingBuoys %>% group_by(BuoyNumber) %>%
      summarize_each(funs(mean(., na.rm=TRUE)))
    
    matchingBuoys <- as.data.frame(matchingBuoys)
    
    #function to calculate weighted averages
    weightedByDistance <- function(x){
      return(weighted.mean(x, w= 1/(matchingBuoys$distToStorm), na.rm = TRUE))
    }
    
    #calculates weighted avgs of buoy variables for each row
    meansForRow <- apply(matchingBuoys[,c(6:18)], 2, weightedByDistance)
    
    #if weighted avgs not all NaN then add on corresponding storm rowNum
    if(sum(meansForRow, na.rm=TRUE) > 0){
      meansForRow <- c(meansForRow, k)
      names(meansForRow)[length(meansForRow)] <- "rowNum"
      meansForRow <- as.data.frame(t(as.matrix(meansForRow)))
      
        #make a dataframe of weighted means for each buoy variable with correspond storm rowNum
        if(length(weightedMeans)){
        weightedMeans <- meansForRow
        
      } else {
        weightedMeans <- rbind(weightedMeans, 
                             meansForRow)
      }
    }
    }
    }
}

```


PROBLEMS TO ADDRESS:
1. need to change 2e7 later to distance we actually want buoys to be from storm
2.  need to fix averaging to deal w repeated rows or repeated parts of row ie 9 and 10 in this example read in 
3. need to figure out what to do w 999 and 99 as missing data (convert to NA that is removed in rm.na)?
4. Check over formatting -- seems like loops are not indented properly 

